VarCovar Rembrandt

### 2023-09-21 reporting/simu_rembrandt_breast.Rmd
update with new simu

- number of genes : $NG = \{10, 50, 200\}$
- proportion of significant genes : $p_g = \{0.2, 0.5\}$
- Variance : $C_{jj} = 0.2$
- Type (D) : $\beta \sim \mathcal{N}(0, 0.4^2)$
- Type (E) : half of genes follows $\beta \sim \mathcal{N}(-0.4, 0.2^2)$ and half follows $\beta \sim \mathcal{N}(0.4, 0.2^2)$
- Type (F) : half of genes follows $\beta \sim \mathcal{N}(-0.8, 0.4^2)$ and half follows $\beta \sim \mathcal{N}(0.8, 0.4^2)$
- Case (IV) : Overall correlation follows a non-standard beta $corr \sim NSBeta(20, 20, min=-1, max =1)$
- Case (V) : Correlation between significant genes $corr \sim NSBeta(10, 10, min=-1, max =1)$ and between other genes $corr \sim NSBeta(25, 25, min=-1, max =1)$

Nb of simulations to perform = $3*2*3*2 = 36$

@@

>>>>

- number of genes : $NG = \{10, 50, 200\}$
- number of individuals : $N = \{50, 100\}$
- proportion of significant genes : $p_g = \{0.2, 0.5\}$
- proportion of censored genes : $c = \{0, 0.3\}$
- Variance : $C_{jj} = 0.2$

- Type (D) : $\beta \sim \mathcal{N}(0, 0.4^2)$
- Type (E) : half of genes follows $\beta \sim \mathcal{N}(-0.4, 0.2^2)$ and half follows $\beta \sim \mathcal{N}(0.4, 0.2^2)$
- Type (F) : half of genes follows $\beta \sim \mathcal{N}(-0.8, 0.4^2)$ and half follows $\beta \sim \mathcal{N}(0.8, 0.4^2)$

**Note : scenario D, E and F do not rely on correlation matrix to simulate variance covariance matrix so the those matrix are unlikely to be positive definite.**

- Type (G), (H) and (I) : same as type (D), (E), (F) where covariance matrix of $\beta$ is obtained from the correlation matrix of gene expression defined by "Case" (instead of no correlation).

- Case (IV) : Overall correlation follows a non-standard beta $corr \sim NSBeta(20, 20, min=-1, max =1)$
- Case (V) : Correlation between significant genes $corr \sim NSBeta(10, 10, min=-1, max =1)$ and between other genes $corr \sim NSBeta(25, 25, min=-1, max =1)$

**Note : Case IV and V are not guaranteed to generate positive definite matrix, after the correlation values are sampled, if the matrix is not positive definite, the nearest positive definite matrix is computed using the algorithm defined by Higham 2002 (see Matrix::nearPD).**

- Case (V) : Correlation between significant genes is $0.2$, correlation with non significant genes is $0$.

@@

**Note : see methods section, the covariance matrix are unlikely to be positive definite.**

```{r fig.height=10, fig.cap="Statistical power depending on simulation setting (second batch). N is the number of patients and NG the number of genes in the pathway."}
source("script/script_simulation/figures_simu_rembrandt_breast_v2.R")
